Data Munging

### 2025-03-03 manuscript/manuscript_file.Rmd
Updating repo to include only files and directories needed to generate results and manuscript

```{r source_scripts, message=FALSE, warning=FALSE, include=FALSE}
# Load results objects from disk for knitting (if they already exist) ----------
# vector of object names
objs <- c("dg",
          "national_stats",
          "ecoreg_stats",
          "em_min_max",
          "FluxBySize",
          "ipcc.n2o"
          )

@@

# list of file paths for objects saved
f <- list()
for(i in 1:N){
    f[i] <- paste0(Sys.getenv("RENV_PROJECT"), "/manuscript/manuscript_files/",
    objs[i], ".rda")
    }

@@

# list of file paths for objects saved
f <- list()
for(i in 1:N){
    f[i] <- paste0(Sys.getenv("RENV_PROJECT"), "/manuscript/manuscript_figures/",
    objs[i])
    }

>>>>

```{r source_scripts, message=FALSE, warning=FALSE, include=FALSE}
# Load results objects from disk for knitting (if they already exist) ----------
# vector of object names
objs <- c("dg",
          "national_stats",
          "ecoreg_stats",
          "em_min_max",
          "FluxBySize"
          )

@@

# list of file paths for objects saved
f <- list()
for(i in 1:N){
    f[i] <- paste0("/manuscript/manuscript_files/",
    objs[i], ".rda")
    }

@@

# list of file paths for objects saved
f <- list()
for(i in 1:N){
    f[i] <- paste0("/manuscript/manuscript_figures/",
    objs[i])
    }

### 2025-03-03 scripts/DG_sensitivity_to_measurement_error.Rmd
Updating repo to include only files and directories needed to generate results and manuscript

```{r setup, warning=FALSE, message=FALSE}
#library(devtools) # needed to install USAboundaries package from
library(knitr)
library(rmarkdown)
library(bayesplot)
library(kableExtra)
library(ggpubr)
library(tidyverse)
library(tidybayes)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(readxl)
library(janitor)
library(sf)

# Install USAboundaries package from GitHub
#devtools::install_github("ropensci/USAboundaries")

@@

## Calculate $\text{N}_2\text{O}$
Below is a summary of the empirical measurements of N2O (ppm) in the
equilibrated headspace.
```{r summary_empirical, echo=TRUE}
load(paste0(localPath, "\\Environmental Protection Agency (EPA)",
            "\\ORD NLA17 Dissolved Gas - Documents\\",
            "inputData\\dg.2021-02-01.RData"))

@@

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

@@

>>>>

```{r setup, warning=FALSE, message=FALSE}
library(knitr)
library(rmarkdown)
library(bayesplot)
library(kableExtra)
library(ggpubr)
library(tidyverse)
library(tidybayes)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(readxl)
library(janitor)
library(sf)
library(usmap)

@@

## Calculate $\text{N}_2\text{O}$
Below is a summary of the empirical measurements of N2O (ppm) in the
equilibrated headspace.
```{r summary_empirical, echo=TRUE}
load("./../inputData/dg.RData")

@@

## Load state boundaries----
states <- usmap::us_map() %>%
  dplyr::filter(!full %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  sf::st_transform(5070) # convert to CONUS Albers

@@

# Update data with uncertainty info
Finally, we update the dg data frame to include information on estimated measurement error

``` {r update_dg_data}
dg <- dg %>%
  mutate(# recompute using error from uncertainty simulations
         n2o.src.snk.error = case_when(
           abs(n2o.sat.ratio - 1) < quantile(S_a_me_abs, probs = c(0.975)) ~ "undetermined",
           TRUE ~ n2o.src.snk))

save(dg, file = "./../inputData/dg.RData")
```

# Session Info

```{r session}
sessionInfo()
```

### 2025-03-03 scripts/NLA_N2O_models.Rmd
Updating repo to include only files and directories needed to generate results and manuscript

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
library(ggpubr)
library(moments)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(kableExtra)
library(ggrepel)
library(dplyr)
library(tidyverse)
library(tidyr)
library(future)
library(foreach)
library(itertools)
library(bayesplot)
library(tidybayes)
library(brms)
library(tictoc)
library(arrow)

@@

# Load model and data objects from drive for knitting (if they already exist)
## vector of object names
objs <- c("dg",
          "df_model",
          "df_test",
          "sframe",
          "pframe",
          "samp_props",
          "n2o_mod1",
          "n2o_mod2",
          "n2o_mod3",
          "n2o_mod4",
          "n2o_mod5",
          "n2o_mod6",
          #"all_predictions_b",
          "n2o_survey_ests",
          "n2o_survey_ests_size",
          "sat_survey_ests",
          "sat_survey_ests_size"
          )

@@

# list of file paths for objects saved
f <- list()
for(i in 1:N){
    f[i] <- paste0(Sys.getenv("RENV_PROJECT"), "/modelFiles/",
    objs[i], ".rda")
    }

@@

That originally munged gas dataset was imported below.

```{r import_data, eval=FALSE, include=TRUE}
load( file = paste0( localPath,
              "/Environmental Protection Agency (EPA)/",
              "ORD NLA17 Dissolved Gas - Documents/",
              "inputData/dg.2021-02-01.RData")
      )

save(dg, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/dg.rda") 
```

@@

save(df_model, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_model.rda") 
```

@@

save(df_test, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/df_test.rda") 
```

@@

```{r import_sample_frame, eval=FALSE, include=TRUE}
df_pop <- read.csv(file = paste0(localPath,
              "/Environmental Protection Agency (EPA)/",
              "ORD NLA17 Dissolved Gas - Documents/",
              "inputData/NLA_Sample_Frame.csv"), header = T)

@@

Below, National mean estimates for dissolved N2O from the model and
design-based approaches were compared. The sample-based estimate was
also included as a reference. The black, vertical, dashed line indicates
the mean of the sample.

```{r n2o_means_national, message=FALSE, cache=TRUE}
all_predictions %>%
  group_by(.draw) %>%
  summarise(mean_n2o = mean(n2o)) %>%
  summarise(estimate = round(median(mean_n2o), 2), # posterior median
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2)) %>% 
  mutate(type = "model") %>%
  bind_rows(cbind(n2o_survey_ests[10, 2:4], type = rep("survey", 1))) %>%
  add_row(estimate = round(mean(df_model$n2o), 2),
          type = "sample") %>%
  print()
```

```{r plot_n2o_means_national, echo=FALSE, fig.align='center', fig.asp=0.5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
all_predictions %>%
  group_by(.draw) %>%
  summarise(mean_n2o = mean(n2o)) %>%
  summarise(estimate = round(median(mean_n2o), 2), # posterior median
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2)) %>% 
  mutate(type = "model") %>%
  bind_rows(cbind(n2o_survey_ests[10, 2:4], type = rep("survey", 1))) %>%
  mutate(cl_width = round(UCL - LCL, 2)) %>%
  ggplot(aes(x = type, y = estimate, color = type)) +
  geom_point(size = 2, position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  geom_hline(yintercept = round(mean(df_model$n2o), 2), 
             linetype = "dashed",
             color = "black") +
  ylab("mean N2O concentration") +
  ggtitle("National estimate comparison") +
  coord_flip() + 
  theme_tidybayes()
```

@@

```{r n2o_mean_wsa9, message=FALSE, warning=FALSE, cache=TRUE}
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(mean_n2o = mean(n2o)) %>%
  group_by(WSA9, .groups = "drop") %>%
  summarise(estimate = round(median(mean_n2o), 2),
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2),
    .groups = "drop") %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(type = "model") %>%
  select(ecoregion, estimate, LCL, UCL, type) %>%
  mutate(ecoregion = forcats::fct_reorder(ecoregion, estimate)) %>%
  bind_rows(cbind(n2o_survey_ests[-10,], type = rep("survey", 9))) %>%
  arrange(ecoregion) %>%
  print()
```

```{r plot_mean_n2o_wsa9, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(mean_n2o = mean(n2o), .groups = "drop") %>%
  group_by(WSA9) %>%
  summarise(estimate = round(median(mean_n2o), 2),
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2),
    .groups = "drop") %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(type = "model") %>%
  select(ecoregion, estimate, LCL, UCL, type) %>%
  mutate(ecoregion = forcats::fct_reorder(ecoregion, estimate)) %>%
  bind_rows(cbind(n2o_survey_ests[-10,], type = rep("survey", 9))) %>%
  ggplot(aes(x = ecoregion, y = estimate, group = type, color = type)) +
  geom_point(size = 2, position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  coord_flip() + 
  theme_tidybayes() +
  ylab("mean dissolved N2O") +
  ggtitle("Ecoregion estimates comparison")
```

@@

```{r table_size_mean_n2o, echo=FALSE,message=FALSE, warning=FALSE, cache=TRUE}
all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_n2o = mean(n2o), .groups = "drop") %>%
  group_by(size_cat) %>%
  summarise(estimate = round(median(mean_n2o), 1),
    LCL = round(quantile(mean_n2o, probs = 0.025), 1),
    UCL = round(quantile(mean_n2o, probs = 0.975), 1),
    .groups = "drop") %>%
  mutate(type = "model") %>%
  bind_rows(cbind(n2o_survey_ests_size, type = rep("survey", 5))) %>%
  mutate(size = factor(size_cat)) %>%
  mutate(size = forcats::fct_reorder(size, estimate)) %>%
  mutate(cl_width = UCL - LCL) %>%
  arrange(size) %>%
  select(size, estimate, LCL, UCL, type) %>% 
  print()
```

```{r plot_size_mean_n2o, echo=FALSE, fig.align='center', fig.asp=1.5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_n2o = mean(n2o), .groups = "drop") %>%
  group_by(size_cat) %>%
  summarise(estimate = round(median(mean_n2o), 2),
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2),
    .groups = "drop") %>% 
  #mutate(size_cat = factor(size_cat)) %>%
  mutate(type = "model") %>%
  select(size_cat, estimate, LCL, UCL, type) %>%
  bind_rows(cbind(n2o_survey_ests_size, type = rep("survey", 5))) %>%
  mutate(size_cat = forcats::fct_reorder(size_cat, estimate)) %>%
  ggplot(aes(x = size_cat, y = estimate, group = type, color = type)) +
  geom_point(size = 2, position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  coord_flip() + 
  theme_tidybayes() +
  ylab("mean dissolved N2O") +
  ggtitle("Size category estimates comparison")
```

>>>>

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
library(ggpubr)
library(moments)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(kableExtra)
library(ggrepel)
library(dplyr)
library(tidyverse)
library(tidyr)
library(future)
library(foreach)
library(itertools)
library(bayesplot)
library(tidybayes)
library(brms)
library(tictoc)
library(arrow)
library(archive)

@@

# Load model and data objects from drive for knitting (if they already exist)
## vector of object names
objs <- c("dg",
          "df_model",
          "df_test",
          "sframe",
          "pframe",
          "samp_props",
          "n2o_mod1",
          "n2o_mod2",
          "n2o_mod3",
          "n2o_mod4",
          "n2o_mod5",
          "n2o_mod6",
          "all_predictions_b"
          )

@@

# list of file paths for objects saved
f <- list()
for(i in 1:N){
    f[i] <- paste0("./../modelFiles/",
    objs[i], ".rda")
    }

@@

That originally munged gas dataset is imported below.

```{r import_data, eval=FALSE, include=TRUE}
load("./inputData/dg.RData")
```

@@

save(df_model, file = "/modelFiles/df_model.rda") 
```

@@

save(df_test, file = "modelFiles/df_test.rda") 
```

@@

```{r import_sample_frame, eval=FALSE, include=TRUE}
df_pop <- read.csv(archive::archive_read("inputData/NLA_Sample_Frame.7z")) # This takes a bit

@@

Below, National mean estimates for dissolved N2O from the model and
design-based approaches were compared. The code and results for the 
design-based approach are located at: 
https://github.com/USEPA/DissolvedGasNla/tree/master/scripts

The sample-based estimates were also included as a reference and are
represented by the black, vertical, dashed lines.

```{r import_design_ests, message=FALSE, cache=TRUE}
### Dissolved N2O estimates

n2o_survey_ests <- read.csv(file = "./../inputData/populationEstimates/Survey_Ests_N20_dissolved.csv") %>%
  tibble() %>%
  rename(estimate = Estimate,
         LCL = LCB95Pct,
         UCL = UCB95Pct) %>%
  select(Type, Subpopulation, estimate, LCL, UCL)

# ests by subpopulations
n2o_survey_ecoregion <- n2o_survey_ests %>%
  filter(Type == "AG_ECO9_NM") %>%
  rename(ecoregion = Subpopulation) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = recode(ecoregion,
                            "Coastal Plains" = "CPL",
                            "Northern Appalachians" = "NAP",
                            "Northern Plains" = "NPL",
                            "Southern Appalachians" = "SAP",
                            "Southern Plains" = "SPL",
                            "Temperate Plains" = "TPL",
                            "Upper Midwest" = "UMW",
                            "Western Mountains" = "WMT",
                            "Xeric" = "XER"))

n2o_survey_size <- n2o_survey_ests %>%
  filter(Type == "AREA_CAT6") %>%
  rename(size_cat = Subpopulation) %>%
  select(size_cat, estimate, LCL, UCL)

### Saturation ratio estimates
sat_survey_ests <- read.csv(file = "./../inputData/populationEstimates/Survey_Ests_N20_sat.csv") %>%
  tibble() %>%
  rename(estimate = Estimate,
         LCL = LCB95Pct,
         UCL = UCB95Pct) %>%
  select(Type, Subpopulation, estimate, LCL, UCL)
```

```{r n2o_means_national, message=FALSE, cache=TRUE}
all_predictions %>%
  group_by(.draw) %>%
  summarise(mean_n2o = mean(n2o)) %>%
  summarise(estimate = round(median(mean_n2o), 2), # posterior median
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2)) %>% 
  mutate(type = "model") %>%
  bind_rows(cbind(n2o_survey_ests[1, 3:5], type = rep("survey", 1))) %>%
  add_row(estimate = round(mean(df_model$n2o), 2),
          type = "sample") %>%
  print()
```

```{r plot_n2o_means_national, echo=FALSE, fig.align='center', fig.asp=0.5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
all_predictions %>%
  group_by(.draw) %>%
  summarise(mean_n2o = mean(n2o)) %>%
  summarise(estimate = round(median(mean_n2o), 2), # posterior median
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2)) %>% 
  mutate(type = "model") %>%
  bind_rows(cbind(n2o_survey_ests[1, 3:5], type = rep("survey", 1))) %>%
  mutate(cl_width = round(UCL - LCL, 2)) %>%
  ggplot(aes(x = type, y = estimate, color = type)) +
  geom_point(size = 2, position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  geom_hline(yintercept = round(mean(df_model$n2o), 2), 
             linetype = "dashed",
             color = "black") +
  ylab("mean N2O concentration") +
  ggtitle("National estimate comparison") +
  coord_flip() + 
  theme_tidybayes()
```

@@

```{r n2o_mean_wsa9, message=FALSE, warning=FALSE, cache=TRUE}
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(mean_n2o = mean(n2o)) %>%
  group_by(WSA9, .groups = "drop") %>%
  summarise(estimate = round(median(mean_n2o), 2),
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2),
    .groups = "drop") %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(type = "model") %>%
  select(ecoregion, estimate, LCL, UCL, type) %>%
  mutate(ecoregion = forcats::fct_reorder(ecoregion, estimate)) %>%
  bind_rows(cbind(n2o_survey_ecoregion, type = rep("survey", 9))) %>%
  #bind_rows(cbind(n2o_survey_ests[-10,], type = rep("survey", 9))) %>%
  arrange(ecoregion) %>%
  print()
```

```{r plot_mean_n2o_wsa9, echo=FALSE, fig.align='center', fig.asp=2, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(mean_n2o = mean(n2o), .groups = "drop") %>%
  group_by(WSA9) %>%
  summarise(estimate = round(median(mean_n2o), 2),
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2),
    .groups = "drop") %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(type = "model") %>%
  select(ecoregion, estimate, LCL, UCL, type) %>%
  mutate(ecoregion = forcats::fct_reorder(ecoregion, estimate)) %>%
  bind_rows(cbind(n2o_survey_ecoregion, type = rep("survey", 9))) %>%
  ggplot(aes(x = ecoregion, y = estimate, group = type, color = type)) +
  geom_point(size = 2, position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  coord_flip() + 
  theme_tidybayes() +
  ylab("mean dissolved N2O") +
  ggtitle("Ecoregion estimates comparison")
```

@@

```{r table_size_mean_n2o, echo=FALSE,message=FALSE, warning=FALSE, cache=TRUE}
all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_n2o = mean(n2o), .groups = "drop") %>%
  group_by(size_cat) %>%
  summarise(estimate = round(median(mean_n2o), 1),
    LCL = round(quantile(mean_n2o, probs = 0.025), 1),
    UCL = round(quantile(mean_n2o, probs = 0.975), 1),
    .groups = "drop") %>%
  mutate(type = "model") %>%
  bind_rows(cbind(n2o_survey_size, type = rep("survey", 5))) %>%
  mutate(size = factor(size_cat)) %>%
  mutate(size = forcats::fct_reorder(size, estimate)) %>%
  mutate(cl_width = UCL - LCL) %>%
  arrange(size) %>%
  select(size, estimate, LCL, UCL, type) %>% 
  print()
```

```{r plot_size_mean_n2o, echo=FALSE, fig.align='center', fig.asp=1.5, fig.width=5, message=FALSE, warning=FALSE, cache=TRUE}
all_predictions %>%
  group_by(size_cat, .draw) %>%
  summarise(mean_n2o = mean(n2o), .groups = "drop") %>%
  group_by(size_cat) %>%
  summarise(estimate = round(median(mean_n2o), 2),
    LCL = round(quantile(mean_n2o, probs = 0.025), 2),
    UCL = round(quantile(mean_n2o, probs = 0.975), 2),
    .groups = "drop") %>% 
  #mutate(size_cat = factor(size_cat)) %>%
  mutate(type = "model") %>%
  select(size_cat, estimate, LCL, UCL, type) %>%
  bind_rows(cbind(n2o_survey_size, type = rep("survey", 5))) %>%
  mutate(size_cat = forcats::fct_reorder(size_cat, estimate)) %>%
  ggplot(aes(x = size_cat, y = estimate, group = type, color = type)) +
  geom_point(size = 2, position=position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = LCL, ymax = UCL) , position=position_dodge(width=0.5)) +
  scale_colour_manual(values = c("black", "grey")) +
  coord_flip() + 
  theme_tidybayes() +
  ylab("mean dissolved N2O") +
  ggtitle("Size category estimates comparison")
```

### 2025-03-03 scripts/dataMunge.Rmd
Updating repo to include only files and directories needed to generate results and manuscript

---
title: "Data Munging"
authors: "J. Beaulieu, R. Martin, and M. McManus"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    fig_caption: yes
    depth: 2
    number_sections: true
editor_options: 
  chunk_output_type: console
---
```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
# load libraries
library(tidyverse) # dplyr, ggplot

@@

# Purpose
The purpose of this .rmd is to prepare the data set for subsequent modeling.  The data will be written as .RData object and stored at the shared documents library associated with this project: https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas

The dissolved gas data were curated in the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2021-01-13.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.

@@

```{r}
# Read dissolved gas data file
dg <- read.table(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/nla17gasDataAggregated_2021-02-01.txt"),
                 header = TRUE, sep = "\t", as.is = TRUE) %>%
  filter(!grepl("AK", site.id)) # omit Alaska sites
```

## Manipulate
### Implement unit conversion and create source/sink column:
``` {r}
dg <- dg %>%
  # unit conversion
  mutate(dissolved.ch4.umol = dissolved.ch4 * 1000000, # mol/L -> umol/L
         dissolved.co2.umol = dissolved.co2 * 1000000, # mol/L -> umol/L,
         dissolved.n2o.nmol = dissolved.n2o * 1000000000, # mol/L -> nmol/L
         sat.ch4.umol = sat.ch4 * 1000000, # mol/L -> umol/L
         sat.co2.umol = sat.co2 * 1000000, # mol/L -> umol/L
         sat.n2o.nmol =  sat.n2o * 1000000000, # mol/L -> nmol/L
         # add source/sink column
         co2.src.snk = ifelse(co2.sat.ratio > 1, "source", "sink"),
         ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink")) %>%
  # remove fields no longer needed
  select(-dissolved.ch4, -dissolved.co2, -dissolved.n2o,
         -sat.ch4, -sat.co2, - sat.n2o)
```

@@

```{r}
dg <- dg %>% 
  mutate(
    # e.ch4 umol CH4/m2/d 
    e.ch4.umol.d = ((dissolved.ch4.umol - sat.ch4.umol) * 1000) * # 1000 L to m3
           (k600.day * (24/100)), # 24 hour to day.  100 cm to m
    # e.co2 umol CO2/m2/d
    e.co2.umol.d = ((dissolved.co2.umol - sat.co2.umol) * 1000) * # 1000 L to m3
           (k600.day * (24/100)), # 24 hour to day.  100 cm to m
        # e.n2o nmol N2O/m2/d
    e.n2o.nmol.d = ((dissolved.n2o.nmol - sat.n2o.nmol) * 1000) * # 1000 L to m3
           (k600.day * (24/100)), # 24 hour to day.  100 cm to m
    # total flux of CH4 per day kmol CH4 day-1
    f.ch4.km.d = e.ch4.umol.d * area.ha * 10000 * (1/10^9), # 1ha = 10,000 m2. 10^9 umol->mmol->mol->kmol
    # total flux of CO2 per day kmol CO2 day-1
    f.co2.km.d = e.co2.umol.d * area.ha * 10000 * (1/10^9), # 1ha = 10,000 m2. 10^9 umol->mmol->mol->kmol
    # total flux of N2O per day mol N2O day-1
    f.n2o.m.d = e.n2o.nmol.d * area.ha * 10000 * (1/10^9))  # 1ha = 10,000 m2. 10^9 nmol->umol->mmol->mol

@@

## Emission rate for population estimate
Tom Kincaid created a file to be used for population estimates.  Lets add our emission rates estimates to this file.  We can then use Tom's code to construct population estimates of emission rates.
```{r}
cdf <- read.csv(paste0(localPath,
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/populationEstimates/",
            "NLA17_DissolvedGases_forPopEst.csv"))

# although data I provided Tom (nla17gasAggregatedForPopEst2019-10-23.txt) included data from both visits,
# data used for population estimates included only visit one.  I don't know rationale for this, but will
# do the same with the revised file here.  Also need to strip out air samples.
unique(cdf$VISIT_NO) # 1
dg.1 <- dg %>% filter(visit.no == 1, sample.source == "DG")

# even after filtering, dg has 85 more observations than cdf
dim(cdf) # 1005, 81
dim(dg.1) # 1090, 306

# which observations in filtered dg, not in cdf?
# many of these are handpicked (HP) sites.  Others not sure
# why are excluded
dg.1 %>% filter(!(dg.1$site.id %in% cdf$SITE_ID)) %>%
  select(site.id) %>%
  as_tibble() %>% print(n=Inf)

# using left_join to exclude sites from dg.1 that Tom did
# not include in cdf (see above).  Presumably the excluded sites
# are not probability sites
cdf.dg <- left_join(cdf, 
                    select(dg.1, site.id, visit.no, 
                           e.ch4.umol.d , e.co2.umol.d, e.n2o.nmol.d, 
                           f.ch4.km.d, f.co2.km.d, f.n2o.m.d) %>%
                      # rename for consistency with Tom's data
                      rename_all(~stringr::str_to_upper(.)) %>% 
                      rename_all(~stringr::str_replace_all(., '\\.', '_')))

dim(cdf.dg) # 1005, 87 same as cdf above, so good

write.csv(cdf.dg, 
          paste0(localPath,
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/populationEstimates/",
            "NLA17_DissolvedGases_forPopEst.jb.csv"))

## Write out data
Write out final .RData object.
```{r}
save(dg, file = paste0(localPath,
                       "/Environmental Protection Agency (EPA)/",
                       "ORD NLA17 Dissolved Gas - Documents/",
                       "inputData/dg.", Sys.Date(), ".RData"))

>>>>

---
title: "Data Munging"
authors: "J. Beaulieu, R. Martin, and M. McManus"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 5
    number_sections: true
    keep_html: true
editor_options: 
  chunk_output_type: console
---
```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
# load libraries
library(tidyverse) # dplyr, ggplot

@@

# Purpose
The purpose of this .rmd was to prepare the data set for subsequent modeling. The data were written as .RData object that is stored in the repository associated with this project: https://github.com/USEPA/DissolvedGasNla

The dissolved gas data were originally curated in the 'NLA' dissolved gas project in RStudio. The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2021-01-13.txt, which is also avaialble in the GitHub repository for this project: https://github.com/USEPA/DissolvedGasNla

The data object ('dg.Rdata') munged in this document was subsequently used to estimate measurement uncertainty in the N2O observations, which is documented in the 'DG_sensitivity_to_measurement_error.Rmd' (Saturation ratio/Standard GC sections) file that is also available in the repository. From there, an updated 'dg.RData' object including this new information on uncertainty was imported into additional RStudio document workflows for (1) making design-based estimates from the survey data (Design_Based_Estimates.Rmd) and (2) generating population estimates using a Bayesian hierarchical model (NLA_N2O_models.Rmd). The object is also imported for the files used to generate the manuscript (manuscript_file.Rmd and manuscript_support.Rmd). All of these files (and more) are included in the GitHub repository for this project in order to ensure the results were reproducible starting from this base file of the observed data.

@@

```{r read_data}
# Read dissolved gas data file
dg <- read.table(file = "./../inputData/nla17gasDataAggregated_2021-02-01.txt",
                 header = TRUE, sep = "\t", as.is = TRUE) %>%
  filter(!grepl("AK", site.id)) # omit Alaska sites

# Read simulated saturation ratio values (See DG_sensitivity_to_measurement_error.Rmd, Saturation ratio/Standard GC)
#S_a_me_abs <- readRDS("inputData/S_a_me_abs.RDS")
```

## Manipulate
### Implement unit conversion and create source/sink column:
``` {r mutate_data}
dg <- dg %>%
  # unit conversion
  mutate(dissolved.ch4.umol = dissolved.ch4 * 1000000, # mol/L -> umol/L
         dissolved.co2.umol = dissolved.co2 * 1000000, # mol/L -> umol/L,
         dissolved.n2o.nmol = dissolved.n2o * 1000000000, # mol/L -> nmol/L
         sat.ch4.umol = sat.ch4 * 1000000, # mol/L -> umol/L
         sat.co2.umol = sat.co2 * 1000000, # mol/L -> umol/L
         sat.n2o.nmol =  sat.n2o * 1000000000, # mol/L -> nmol/L
         # add source/sink column
         co2.src.snk = ifelse(co2.sat.ratio > 1, "source", "sink"),
         ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink")
         ) %>%
  # remove fields no longer needed
  select(-dissolved.ch4, -dissolved.co2, -dissolved.n2o,
         -sat.ch4, -sat.co2, - sat.n2o)
```

@@

```{r mutate_data_2}
dg <- dg %>% 
  mutate(
    # e.ch4 umol CH4/m2/d 
    e.ch4.umol.d = ((dissolved.ch4.umol - sat.ch4.umol) * 1000) * # 1000 L to m3
           (k600.day * (24/100)), # 24 hour to day.  100 cm to m
    # e.co2 umol CO2/m2/d
    e.co2.umol.d = ((dissolved.co2.umol - sat.co2.umol) * 1000) * # 1000 L to m3
           (k600.day * (24/100)), # 24 hour to day.  100 cm to m
        # e.n2o nmol N2O/m2/d
    e.n2o.nmol.d = ((dissolved.n2o.nmol - sat.n2o.nmol) * 1000) * # 1000 L to m3
           (k600.day * (24/100)), # 24 hour to day.  100 cm to m
    # total flux of CH4 per day kmol CH4 day-1
    f.ch4.km.d = e.ch4.umol.d * area.ha * 10000 * (1/10^9), # 1ha = 10,000 m2. 10^9 umol->mmol->mol->kmol
    # total flux of CO2 per day kmol CO2 day-1
    f.co2.km.d = e.co2.umol.d * area.ha * 10000 * (1/10^9), # 1ha = 10,000 m2. 10^9 umol->mmol->mol->kmol
    # total flux of N2O per day mol N2O day-1
    f.n2o.m.d = e.n2o.nmol.d * area.ha * 10000 * (1/10^9))  # 1ha = 10,000 m2. 10^9 nmol->umol->mmol->mol

@@

## Write out data
Write out final .RData object.
```{r save_data_munged}
save(dg, file = "./../inputData/dg.RData")
```
